Synthesizing Symmetric Lenses

### 2018-07-10 papers/SymmetricLenses/main.tex
in progress edits to intro

\subsection{Optimizations}
\label{subsec:optimizations}

@@

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented all Boomerang's asymmetric lens combinators using a combination of
the symmetric lens combinators, as well as other symmetric lens combinators for
the asymmetric theory extensions present in Boomerang, like matching lenses and
quotient lenses. We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
alongside lens combinators, use those synthesis results in the manual writing of
new lenses, and use previously defined lenses in the synthesis of new lenses.

@@

\item How much slower is our tool on bijective lens synthesis benchmarks than
  the existing Optician system customized for bijective lenses?

@@

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted \afm{8} data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
\item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
  ad-hoc data present in Linux configuration files into a structured, dictionary
  representation.
\item We used \afm{8} benchmarks from synchronization tasks between a
  variety of real-life formats. These tasks range from synchronizing REST
  and JSON web resource descriptions to synchronizing BibTeX and EndNote
  citation descriptions.
\end{enumerate}

@@

To begin, we measure the run time of our algorithm in 2 modes:

\begin{tabulary}{\linewidth}{rL}
  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
  \SSNCOpt{}: & The symmetric synthesis algorithm, but compositional synthesis not enabled.\\
\end{tabulary}\\

Recall that compositional synthesis exploits user annotations to break a benchmark into a series of smaller synthesis steps. In our experience, compositional synthesis (SS mode) allows our system to scale to arbitrarily large and complex formats. However, SSNC mode, which uses no such annotations and synthesizes a complete lens all at once provides a useful experimental stress test for the system. Recall from Section

@@

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 30 seconds, and average the result over 5 runs.
Figure~\ref{fig:times} summarizes the results of these tests. We find that our
algorithm is able to synthesize all of the benchmarks in our benchmark suite in
under 30 seconds. Without compositional synthesis, the synthesis algorithm is
able to solve 26 problem instances. Recall from Section~\ref{subsec:optimizations} that users may tune the termination metric. Success on this set of benchmarks requires we tune that termination codition (e.g. for some benchmarks, allowing the system flexibility to spend more time considering more of the search space). In Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing the termination metric.

@@

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmark suite. \afm{8} of these
benchmarks had to be slightly altered with tighter specifications or termination
parameters\dpw{What does that mean? You had to alter it to get the system to work?}, as we had to alter the original specifications to generate the
correct lens.

@@

\begin{tabulary}{\linewidth}{rL}
  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
  \BSOpt{}: & The existing bijective synthesis
              algorithm with all optimizations enabled.\\
\end{tabulary}\\

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 30 seconds, and average the result over 5 runs.
Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
depth 3 into a serialized version of the structured dictionary representation
used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
complete.

\begin{figure}
  \includegraphics{generated-graphs/times_bijective}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \BSOpt{} is the full bijective lens synthesis algorithm.}
  \label{fig:times_bijective}
\end{figure}

\subsection{The Effects of Our Heuristics}
\label{subsec:effect-of-heuristics}
We evaluate the usefulness of (1) our information theoretic metric, and (2) our
termination heuristic.  To this end, we run our program in 3 different modes:\\
\begin{tabulary}{\linewidth}{rL}
  \AnyOpt{}: & We synthesize any lens \dpw{Did you mean the following? But is the order truly arbitrary? Is size or anything else a driver?:] We synthesize lenses in an arbitrary order (i.e. without use of the information theoretic preference metric)} that satisfies the specification, with no
               notion of cost.\\
  \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
              returns that lens.\dpw{What controls the ``greediness''? Say a little something else to more clearly differentiate from Any.}\\
  \CCOpt{}: & We replace our information theoretic cost metric with one where
              the cost of the lens is the number of disonnects plus the number
              of merges.\\
  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
    universally set to 0 (default). \\
  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
    termination parameter universally set to 25. \\
  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
    termination parameter universally set to -25. \\
\end{tabulary}

@@

We experimented with the \CCOpt{} mode to determine whether the complexity of the information theoretic measure was really needed. Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures. The \CCOpt{} mode is an example of such a simple measure --- it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union. Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
our metrics are very important for finding the correct lens. Our
information-theoretic metric is critical for finding the correct lens. Indeed,
only \afm{10} of our benchmarks succeeded when running in \CCOpt{} mode.

>>>>

\subsection{Optimizations}

@@

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented all Boomerang's asymmetric lens combinators using a combination of
the symmetric lens combinators, as well as other symmetric lens combinators for
the asymmetric theory extensions present in Boomerang, like matching lenses and
quotient lenses.

We have extended the Optician synthesis plugin to Boomerang~\cite{?} with our
symmetric synthesis algorithm. In doing so, users can write synthesis tasks
alongside lens combinators, use those synthesis results in the manual writing of
new lenses, and use previously defined lenses in the synthesis of new lenses.

@@

\item How much slower do we perform on bijective lens synthesis benchmarks than
  the existing Optician system customized for bijective lenses?

@@

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In these
  benchmarks, we not only provide functions that clean the data, but we also
  provide other synchronization functions: if the cleaned data is updated, those
  changes are reflected in the original format.
\item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
  ad-hoc data present in Linux configuration files into a structured, dictionary
  representation.
\item We used \afm{8} benchmarks from synchronization tasks between a
  variety of real-life formats. These tasks range from synchronizing REST
  and JSON web resource descriptions to synchronizing BibTeX and EndNote
  citation descriptions.
\end{enumerate}

@@

To answer our first research question, we measure the runtime of our algorithm
on our benchmark suite.

We synthesized lenses in 2 modes:

\begin{tabulary}{\linewidth}{rL}
  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
  \SSNCOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
                but compositional synthesis is not enabled.\\
\end{tabulary}\\

We include running the symmetric synthesis algorithm without compositional
synthesis to stress test the algorithms. While compositional synthesis makes the
procedure easier, and demonstrates the potential to be used in arbitrarily hard
problems, it is dependent on the user identifying important sublenses between
subcomponents of the final formats. It is therefore interesting to see how many
problems each algorithm can solve, when trying to synthesize ``in one go.''

@@

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 30 seconds, and average the result over 5 runs.
Figure~\ref{fig:times} summarizes the results of these tests. We find that our
algorithm is able to synthesize all of the benchmarks in our benchmark suite in
under 30 seconds. Without compositional synthesis, the synthesis algorithm is
able to solve 26 problem instances.

@@

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmark suite. \afm{8} of these
benchmarks had to be slightly altered with tighter specifications or termination
parameters, as we had to alter the original specifications to generate the
correct lens.

@@

\begin{tabulary}{\linewidth}{rL}
  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
  \BSOpt{}: & We synthesize lenses using the existing bijective synthesis
              algorithm with all optimizations enabled.\\
\end{tabulary}\\

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 30 seconds, and average the result over 5 runs.
Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
\BSOpt{}.

\begin{figure}
  \includegraphics{generated-graphs/times_bijective}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
    symmetric synthesis algorithm without using a library of existing lenses.
    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
    existing lenses. Our symmetric synthesis algorithm is able to complete all
    benchmarks in our benchmark suite in under 30 seconds, and is able to
    complete most of the benchmarks without using compositional synthesis.}
  \label{fig:times_bijective}
\end{figure}

The slowest completed benchmark for both synthesis algorithms is
\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
depth 3 into a serialized version of the structured dictionary representation
used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
complete.

\subsection{The Effects of Our Heuristics}
We evaluate the usefulness of (1) our information theoretic metric, and (2) our
termination heuristic.  To this end, we run our program in 3 different modes:\\
\begin{tabulary}{\linewidth}{rL}
  \AnyOpt{}: & We synthesize any lens that satisfies the specification, with no
               notion of cost.\\
  \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
              returns that lens.\\
  \CCOpt{}: & We replace our information theoretic cost metric with one where
              the cost of the lens is the number of disonnects plus the number
              of merges.\\
  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
    universally set to 0 (default). \\
  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
    termination parameter universally set to 25. \\
  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
    termination parameter universally set to -25. \\
\end{tabulary}

@@

Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
our metrics are very important for finding the correct lens. Our
information-theoretic metric is critical for finding the correct lens. Indeed,
only \afm{10} of our benchmarks succeeded when running in \CCOpt{} mode.
